---
title: "Validate Submission"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validate Submission}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `SMHvalidation` R package contains functions to pull data from the R
package covidcast (output in the Scenario Modeling Hub (SMH) standard
format), to validate and visualize Scenario Modeling Hub submissions.

In this vignette, we demonstrate the use of the validation function:
`validate_submision()`. The function can generate 3 different outputs:

-   message when the submission does not contain any issues
-   warning + report message when the submission does contain one or
    multiple minor issues that do not prevent the submission from being
    included.
-   error + report message when the submission contains one or
    multiple minor and/or major issues that prevent the submission from being
    included. In this case the submission file will need to be updated
    to be included in the corresponding SMH round.

We will start by loading `SMHvalidation`:

```{r setup}
library(SMHvalidation)
```

## `validate_submission()`

To run the SMH validation checks, we  use the `validate_submision()`
function. For more information on this function, please consult the
documentation:

```{r eval=FALSE}
?validate_submission()
```

For more information on the tests performed by the function perform, please
consult the vignette [Validation Checks](validation-checks.html).

#### Parameters

The function requires multiple parameters:

-   `path`: path to the submission files. The package contains 3
    examples files that can be used to test the function. There paths
    can be accessed by using:
    `system.file("extdata", "FILENAME", package = "SMHvalidation")`. The
    three filenames (+ quick descriptions) are:
    -   "2022-01-09_team-model.csv" : should returns no issues
    -   "2021-08-15_team-model.gz" : should returns multiple warnings
    -   "2021-11-14_team-model.zip" : should returns multiple warnings
        and errors
-   `js_def`: list containing the round specific and scenario
    information.
-   `lst_gs`: named list of data frame containing the observed data. We
    highly recommend using the output of the `pull_gs_data()` function
    as input. This function will generate the output in the expected
    format with the expected data. For more information, please see
    `?pull_gs_data()`. This parameter can be set to `NULL` is no
    observed data comparison is required.
-   `pop_path`: path to a table containing the population size of each
    geographical entities by FIPS (in a column "location") and by
    location name. For example, path to the locations file in the
    [COVID19 Scenario Modeling Hub GitHub
    repository](https://github.com/midas-network/covid19-scenario-modeling-hub/blob/master/data-locations/locations.csv)

#### Output

The function returns multiple messages and either a warning message or
an error message if the submission contains minor issues and/or major
issues respectively. If the submission contains minor and/or major
issue(s), it will also returns a "report" message containing a detailed
message about the issue and if the issue is:

-   "minor": warning message asking to verify if it's correct but will
    not block the inclusion of the submission to the SMH round.
-   "major": error message asking to fix the issue before including the
    submission in the corresponding SMH round.

#### Prerequisites

Before running the function, we will prepare the required and optional
parameter as explained above:

```{r prerequisite, message=FALSE, warning=FALSE}
# lst_gs <- pull_gs_data() the first week to observed data
lst_gs <- NULL
pop_path <- "https://raw.githubusercontent.com/midas-network/covid19-scenario-modeling-hub/master/data-locations/locations.csv"
js_def <- system.file("extdata", "covid_tasks.json", package = "SMHvalidation")
```

## Example 1: No issues

We can now run the `validate_submission()` function on a first example
file.

```{r test1}
path_test <- system.file("extdata", "2022-01-09_team-model.csv", 
                         package = "SMHvalidation")
validate_submission(path_test, js_def, lst_gs, pop_path)
```

The function returns a simple message stating that no issues has been
found.

## Example 2: Warnings

```{r test2}
path_test <- system.file("extdata", "2021-08-15_team-model.gz", 
                         package = "SMHvalidation")
validate_submission(path_test, js_def, lst_gs, pop_path)
```

The function returns a warning message stating that one or multiple
issues has been found.

The submission will still be accepted but please verify that the
information in the submission is correct.

## Example 3: Warnings & Error

```{r test3, error=TRUE}
path_test <- system.file("extdata", "2021-11-14_team-model.zip", 
                         package = "SMHvalidation")
validate_submission(path_test, js_def, lst_gs, pop_path)
```

The function returns an error message stating that one or multiple
issues has been found.

The submission will NOT be accepted and need correction before being
accepted.

## Example 4: Flu submission

```{r test4}
path_test <- system.file("extdata", "2022-08-14_team-flu.csv", 
                         package = "SMHvalidation")
pop_path <- "https://raw.githubusercontent.com/midas-network/flu-scenario-modeling-hub/main/data-locations/locations.csv"
js_def <- system.file("extdata", "flu_tasks.json", package = "SMHvalidation")
validate_submission(path_test, js_def, NULL, pop_path)
```

## Plotting Projections

The `SMHvalidation` R package also contains plotting functionality to
output a plot of each location and target, with all scenarios and
observed data incorporated (only for quantile type output).

```{r proj_plots}
path_test <- system.file("extdata", "2022-01-09_team-model.csv", 
                         package = "SMHvalidation")
save_path <- getwd()
pdf_path <- file.path(save_path, paste0("2022-01-09", "_", "team-model", 
                                        "_plots.pdf"))
print(paste0("Saving pdf plots to ", pdf_path, "."))
generate_validation_plots(path_proj = path_test, lst_gs = lst_gs, 
                          save_path = getwd(), y_sqrt = FALSE, 
                          plot_quantiles = c(0.025, 0.975))
system(paste0('open "', pdf_path, '"'))
```

Try another file type (CSV, GZ, and ZIP files format are accepted
currently)

```{r proj_plots2}
path_test <- system.file("extdata", "2021-08-15_team-model.gz", 
                         package = "SMHvalidation")
save_path <- getwd()
pdf_path <- file.path(save_path, paste0("2021-08-15", "_", "team-model", 
                                        "_plots.pdf"))
print(paste0("Saving pdf plots to ", pdf_path, "."))
generate_validation_plots(path_proj = path_test, lst_gs = lst_gs, 
                          save_path = getwd(), y_sqrt = FALSE, 
                          plot_quantiles = c(0.025, 0.975))
system(paste0('open "', pdf_path, '"'))
```

Another example is available for FLU submission (as this result is not compared to
observed data, set the `lst_gs` parameter to `NULL`)

```{r proj_plots3}
path_test <- system.file("extdata", "2022-08-14_team-flu.csv", 
                         package = "SMHvalidation")
save_path <- getwd()
pdf_path <- file.path(save_path, paste0("2022-08-14", "_", "team-flu", 
                                        "_plots.pdf"))
print(paste0("Saving pdf plots to ", pdf_path, "."))
generate_validation_plots(path_proj = path_test, lst_gs = NULL, 
                          save_path = getwd(), y_sqrt = FALSE, 
                          plot_quantiles = c(0.025, 0.975))
system(paste0('open "', pdf_path, '"'))
```
