---
title: "Validate Submission"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validate Submission}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `SMHvalidation` R package contains functions to validate and visualize
Scenario Modeling Hub submission, and to pull data from the R
package `covidcast` (output in the Scenario Modeling Hub (SMH) standard
format).

In this vignette, we will show how to use the validation function:
`validate_submision()`. The function can generate 3 different outputs:

-   message when the submission does not contain any issues
-   warning + report message when the submission does contain one or
    multiple minor issues that do not prevent the submission to be
    included.
-   error + report message when the submission does contain one or
    multiple minor and/or major issues that prevent the submission to be
    included. In this case the submission file will have to be updated
    to be include in the
    corresponding SMH round.

We will start by loading `SMHvalidation`:

```{r setup}
library(SMHvalidation)
```

## `validate_submission()`

To run the SMH validation checks, we can use the `validate_submision()`
function. For more information on this function, please consult the
documentation

```{r eval=FALSE}
?validate_submission()
```

For more information on all the tests the function perform, please
consult the vignette [Validation Checks](validation-checks.html).

#### Parameters

The function requires multiple parameters:

-   `path`: path to the submissions file (or folder for partitioned data) to
    test. A vector of parquet files can also be inputted, in this case, the
    validation will be run on the aggregation of all the parquet files
    together, and each file individually should match the expected SMH
    standard. If partition is not set to `NULL`, it should be inputted the path
    to the folder containing the partitioned data.
    The package contains 3 examples files that can be used to test the
    function. There paths can be accessed by using:
    `system.file("extdata", "FILENAME", package = "SMHvalidation")`, the
    3 filenames (+ quick descriptions) are:
    -   `"2022-01-09_team-model.csv"` : should returns no issues
    -   `"2021-08-15_team-model.gz"` : should returns multiple warnings
    -   `"2021-11-14_team-model.parquet"` : should returns multiple warnings
        and errors
-   `js_def`: path to JSON file containing round definitions: names of columns,
    target names, ... following the `tasks.json`
    [Hubverse format](https://hubdocs.readthedocs.io/en/latest/user-guide/hub-config.html#hub-model-task-configuration-tasks-json-file)
    -   the information in the JSON file can be separated in multiple group
        per each round:
        -   The "task_ids" object defines both labels and contents for each
            columns in submission files. Any unique combination of the values
            define a single modeling task. For example, for SMH it can be the
            columns: `"scenario_id"`, `"location"`, `"origin_date"`,
            `"horizon"`, `"target"`, `"age_group"`, `"race_ethnicity"`.
        -   The "output_type" object defines accepted representations for
            each task. For example, for SMH it concerns the columns:
            `"output_type"`, `"output_type_id"`, `"run_grouping"`,
            `"stochastic_run"` and `"value"`.
    
And have some optional parameters:

-   `lst_gs`: named list of data frame containing the observed data. We
    highly recommend to use the output of the `pull_gs_data()` function
    as input. This function will generate the output in the expected
    format with the expected data. For more information, please do
    `?pull_gs_data()`. This parameter can be set to `NULL` (default) to 
     don't compared between the value and the  observed data.
-   `pop_path`: path to a table containing the population size of each
    geographical entities by FIPS (in a column "location") and by
    location name. For example, path to the locations file in the
    [COVID19 Scenario Modeling Hub GitHub
    repository](https://github.com/midas-network/covid19-scenario-modeling-hub/blob/master/data-locations/locations.csv).
     This parameter can be set to `NULL` (default) to don't run a comparison
     between the value and the population size data.
-   `merge_sample_col`: Boolean to indicate if in the submission file(s), the
     output type `"sample"` has the `"output_type_id"` column sets to `NA` and
     the information is contained into 2 columns: `"run_grouping"` and
     `⁠"stochastic_run`⁠". By default, `FALSE`
-   `partition`: character vector indicating if the submission file is
     partitioned and if so, which field (or column) names correspond the path
     segments. By default, `NULL` (no partition). See
     [arrow R package](https://arrow.apache.org/docs/r/) for more information,
     and especially `arrow::write_dataset()`, `arrow::open_dataset()` functions.
     **Warning**: If the submission files is in a "partitioned" format, 
     the `path` parameter should be to a directory to a folder containing 
     **ONLY** the "partitioned" files. If any other file is present in the 
     directory, it will be included in the validation.
-    `n_decimal`: integer, number of decimal point accepted in the column
      `"value"` (only for `"sample"` output type), if `NULL` (default) no limit
      expected.
-    `round_id`: character string, round identifier. This identifier is use to
      extract the associated round information from the `js_def` parameter.
      If `NULL` (default), extracted from path.
      

#### Output

The function returns multiple messages and either a warning message or
an error message if the submission contains minor issues and/or majors
issues respectively. If the submission contains minor and/or major
issue(s), it will also returns a "report" message containing a detailed
message about the issue and if the issue is:

-   "minor": warning message asking to verify if it's correct but will
    not block the inclusion of the submission to the SMH round.
-   "major": error message asking to fix the issue before including the
    submission in the corresponding SMH round.

#### Prerequisite

Before running the function, we will prepare the required and optional
parameter as explained above:

```{r prerequisite, message=FALSE, warning=FALSE}
js_def <- system.file("extdata", "covid_tasks.json", package = "SMHvalidation")
```

## Example 1: No issues

We can now run the `validate_submission()` function on a first example
file.

```{r test1}
path_test <- system.file("extdata", "2022-01-09_team-model.csv",
                         package = "SMHvalidation")
validate_submission(path_test, js_def)
```

The function returns a simple message stating that no issues has been
found.

## Example 2: Warnings

```{r test2}
path_test <- system.file("extdata", "2021-08-15_team-model.gz",
                         package = "SMHvalidation")
validate_submission(path_test, js_def)
```

The function returns a warning message stating that one or multiple
issues has been found.

The submission will still be accepted but please verify that the
information in the submission is correct.

## Example 3: Warnings & Error

```{r test3, error=TRUE}
path_test <- system.file("extdata", "2021-11-14_team-model.parquet",
                         package = "SMHvalidation")
validate_submission(path_test, js_def)
```

The function returns an error message stating that one or multiple
issues has been found.

The submission will NOT be accepted and need correction before being
accepted.

## Plotting Projections

The `SMHvalidation` R package also contains plotting functionality to
output a plot of each location and target, with all scenarios and (optional)
observed data (only for quantile type output).

```{r proj_plots}
path_test <- system.file("extdata", "2022-01-09_team-model.csv",
                         package = "SMHvalidation")
save_path <- getwd()
pdf_path <- file.path(save_path, paste0("2022-01-09", "_", "team-model",
                                        "_plots.pdf"))
print(paste0("Saving pdf plots to ", pdf_path, "."))
generate_validation_plots(path_proj = path_test, lst_gs = NULL,
                          save_path = getwd(), y_sqrt = FALSE,
                          plot_quantiles = c(0.025, 0.975))
system(paste0('open "', pdf_path, '"'))
```

The function accepts multiple file type (CSV, PARQUET (partition and not
partition), GZ, and ZIP files format are accepted currently. 

For more documentation, please consult the function documentation
`?generate_validation_plots`.
