---
title: "Validate Submission"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validate Submission}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `SMHvalidation` R package contains functions to validate and visualize
Scenario Modeling Hub submissions.

In this vignette, we demonstrate the use of the validation function:
`validate_submission()`. The function generates a list representing the
output of the series of checked performed on a specific submission. Each named 
element of the list represents the result of an individual check.

We will start by loading `SMHvalidation`:

```{r setup}
library(SMHvalidation)
```

## `validate_submission()`

To run the SMH validation checks, we can use the `validate_submission()`
function. For more information on this function, please consult the
documentation.

```{r eval=FALSE}
?validate_submission()
```

The function internally runs different validation checks
functions from 
[hubValidations package](https://hubverse-org.github.io/hubValidations/) on
SMH submissions.
For more information on all the tests the function perform, please
consult the vignette [Validation Checks](validation-checks.html).

#### Parameters

The function requires multiple parameters:

-   `path`: path to the submissions file (or folder for partitioned data) 
    to test. PQT, PARQUET, CSV, ZIP (not partitioned), or GZ (not partitioned) 
    file format are accepted. The path should be relative to the path of the hub
    containing the file to check.
-    `hub_path`:  path to the hub containing the submission file.

The package contains an example hub with two example sumbission that can be 
used to test the function. There paths can be accessed by using the 
`system.file()` function. 

```{r setup_path}
hub_path <- system.file("extdata/exp/", package = "SMHvalidation")
```

The example hub follows the hubverse standard and organization. For more 
information on hubverse, please consult the 
[associated website](https://hubverse.io/en/latest/user-guide/hub-structure.html).

It contains example submission for two teams, one partitioned. 
```{r setup_path_f}
path_1 <- "model-output/team2-modelb/2023-11-12-team2-modelb.parquet"
path_2 <- "model-output/t3-mc"
```

###### Optional Parameters
  
The function allows some optional parameters:

-   `js_def`: path to JSON file containing round definitions: names of columns,
    target names, ... following the `tasks.json`
    [Hubverse format](https://hubdocs.readthedocs.io/en/latest/user-guide/hub-config.html#hub-model-task-configuration-tasks-json-file). The version 5.0 minimal is expected. If 
    `NULL` (default), will use the default path `"hub-config/tasks.json"` in 
    the hub using hub_path parameter.
-   `target_data`: path to the target data in the Hubverse time series target 
    data 
    [hubverse standard format](https://hubverse.io/en/latest/user-guide/target-data.html). 
    Set to `NULL `(default), to NOT include comparison with observed data.
-   `pop_path`: path to a table containing the population size (in a column 
    `"population"`) of each  geographical entities by FIPS (in a column 
    `"location"`). Use to compare that value is not higher than expected 
    population size. Set to `NULL` (default), to NOT run comparison with 
    population data.
-   `merge_sample_col`: vector to indicate if the for the output type 
    `"sample"`, the output_type_id column is set to `NA` and the information is 
    contained into other columns. If so the parameter should be set to the 
    sample ID columns names, for example: 
    `⁠c("run_grouping" and "stochastic_run")`⁠.
    By default, `NULL` (sample ID information in the output type id column)
-   `partition`: character vector indicating if the submission file is
    partitioned and if so, which field (or column) names correspond to the path
    segments. By default, `NULL` (no partition). See
    [arrow R package](https://arrow.apache.org/docs/r/) for more information,
    and especially `arrow::write_dataset()`, `arrow::open_dataset()` functions.
    **Warning**: If the submission files is in a "partitioned" format, 
    the `path` parameter should be to a directory to a folder containing 
    **ONLY** the "partitioned" files. If any other file is present in the 
    directory, it will be included in the validation.
-   `n_decimal`: integer, number of decimal points accepted in the column
    `"value"` (only for `"sample"` output type), if `NULL` (default) no limit
    expected.
-   `round_id`: character string, round identifier. If `NULL` (default), 
    extracted from path.
-   `verbose`: Boolean, if TRUE add information about the sample pairing 
    information in output message. By default, TRUE (slows the validation 
    validation for sample output type)
-   `verbose_col`: character vector, details of values included in the pairing
    of each group for specific mentioned columns. If `NULL` (default), no
    additional information in the pairing information. (required `verbose` to be
    set to `TRUE`).
-   `r_schema`: Schema arrow objects, to read partition files with a specific 
    schema. If none provided (default, `NULL`), the schema will be created 
    from the information in the `tasks.json` file in the hub. 

#### Output

The function returns multiple messages and if any issue a failure or 
an error message.

-   if the check succeeds: a message is returned
-   if the check fails: an error or a failure message is returned. A failure 
    represents a check that fails but does not block the proceeding of the 
    validation to other checks. Whereas an error, will stop the validation 
    process. 

## Example 1

We can now run the `validate_submission()` function on a first example
file in the hubverse format.

```{r test1}
validate_submission(path_1, hub_path)
```

## Example 2: Partitioned files

We can now run the `validate_submission()` function on a first example
file in the SMH format, with the sample ID information split into two columns:
`"run_grouping"`, `"stochastic_run"`. For more information on this format,
please consult the 
[Sample file format](https://scenariomodelinghub.org/documentation/sample_format.html)
webpage documentation.

```{r test2}
validate_submission(path_2, hub_path, 
                    partition = c("origin_date", "target"),
                    merge_sample_col = c("run_grouping", "stochastic_run"))
```

## Sample pairing information

The `validate_submission()` function can returns additional information about
the pairing information of the `sample` output type by using the 
`verbose = TRUE` and `verbose_col` parameters. 
The detailed grouping of the value in the `scenario_id` column will be printed 
in the output report:

For example:
```{r test2_age}
validate_submission(path_2, hub_path, 
                    partition = c("origin_date", "target"),
                    merge_sample_col = c("run_grouping", "stochastic_run"),
                    verbose_col = "age_group")
```


## Plotting Projections

The `SMHvalidation` R package also contains plotting functionality to
output a plot of each location and target, with all scenarios and (optional)
observed data (only for quantile type output).

```{r proj_plots, eval=FALSE}
generate_validation_plots(paste0(hub_path, path_1))
```

The function accepts multiple file types (CSV, PARQUET (partition and not
partition), GZ, and ZIP files format are accepted currently). 

For more documentation, please consult the function documentation
`?generate_validation_plots`.
